---
title: "DPA_Project"
author: "Yashwanth Praveen - A20473431 and Brandon Bennitt - A20416733"
date: "11/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(scales)
library(stringr)
library(janitor)
library(stats)
library(rockchalk)
library(chron)
library(ggcorrplot)
library(modelr)
library(caret)
library(olsrr)
library(jtools)
theme_set(theme_light())
```

```{r}
full_data <- read.csv("/Users/yashwanthpraveen/Documents/IIT_3rdSem/Data_Preparation_and_Cleaning/Project/plastic_pollution.csv",header = F)


```

### Working on the column names
```{r}

name_list <- full_data[2,]
data_names <- list(names(full_data))
for (i in 1:ncol(full_data)) {
    names(full_data)[i] <- name_list[i] 
  }
```

### Removing the header overload
```{r}
full_data <- full_data[-c(1,2),]
```

### subsetting our data with required cleaning 
```{r}
pollution_data <- full_data %>% 
  filter(CountryName_FromSource == "United States") %>%
  clean_names() %>%
  type_convert()


```

### Mutating required columns
```{r}
pollution_data <- pollution_data %>% 
  mutate(total_area_sq_m = (total_width_m) * (total_length_m),
         sub_country_l1_from_source = replace(sub_country_l1_from_source,sub_country_l1_from_source == "Or","OR"),
         sub_country_l2_from_source = replace(sub_country_l2_from_source,sub_country_l2_from_source == "Kenai Peninsula Borough","Kenai Peninsula"),
         sub_country_l1_from_source = replace(sub_country_l1_from_source,sub_country_l1_from_source == "Hi","HI"),
         sub_country_l2_from_source = replace(sub_country_l2_from_source,sub_country_l2_from_source == "Hawai'i","Hawaii"),
         sub_country_l1_from_source = replace(sub_country_l1_from_source,sub_country_l1_from_source == "Ha","HI"),
         sub_country_l2_from_source = replace(sub_country_l2_from_source,sub_country_l2_from_source == "Santa Cruz County","Santa Cruz"),
         sub_country_l2_from_source = replace(sub_country_l2_from_source,sub_country_l2_from_source == "Moss Landing","Monterey")) %>%
  filter(sub_country_l2_from_source != "USA") 
  
  


```


```{r}



#county_list <- as.list(unique(pollution_data$sub_country_l2_from_source))
#new_df <- pollution_data %>%
#  select(sub_country_l1_from_source,sub_country_l2_from_source) %>%
#  mutate(country = sub_country_l1_from_source,county = sub_country_l2_from_source) %>%
#  group_by(sub_country_l1_from_source) %>%
#  unique()


#write_csv(new_df,"/Users/yashwanthpraveen/Documents/IIT_3rdSem/Data_Preparation_and_Cleaning/Project/new_df.csv",col_names = T)


```

### Percentage of plastic waste on the beach

```{r}
ppc <- read_csv("/Users/yashwanthpraveen/Documents/IIT_3rdSem/Data_Preparation_and_Cleaning/Project/population_per_county.csv")
merged_data <- merge(x = pollution_data,y = ppc,by = c("sub_country_l1_from_source","sub_country_l2_from_source"),all.x = T)

glimpse(merged_data)
```


### looking into merged_data i.e, What state has the highest garbage to population ratio

```{r}
 q1 <- merged_data %>%
  select(sub_country_l1_from_source,sub_country_l2_from_source,totalltems_event_record,population_2019) %>%
  group_by(sub_country_l1_from_source,sub_country_l2_from_source,population_2019) %>%
  summarise(average = mean(merged_data$totalltems_event_record)) %>%
  mutate(ratio = average/population_2019) %>%
  arrange(desc(ratio))

q1 %>%
  select(average,ratio)
```

### plotting the data
```{r}
q1 %>% 
  head(20) %>%
  ggplot(aes(y = reorder(sub_country_l2_from_source,ratio),x = ratio)) +
  geom_bar(stat = "identity") +
  labs(x = "Ratio",y = "County")
```

### Garbage in each state

```{r}
merged_data %>%
  select(sub_country_l1_from_source,sub_country_l2_from_source,totalltems_event_record,population_2019,political_affiliation) %>%
  mutate(political_affiliation = as.factor(political_affiliation)) %>%
  group_by(sub_country_l1_from_source,political_affiliation) %>%
  arrange(desc(totalltems_event_record)) %>%
  summarise(average = mean(totalltems_event_record)) %>%
  ggplot(aes(y = reorder(sub_country_l1_from_source,average),x = average,fill = political_affiliation)) +
  geom_bar(stat = "identity") +
  labs(x = "State_Average",y = "State") +
  scale_fill_manual("legend", values = c("0" = "blue", "1" = "red"))
```

### Hypothesis Test to know political affiliation

```{r}
pol_data <- merged_data %>%
  select(sub_country_l1_from_source,sub_country_l2_from_source,totalltems_event_record,population_2019,political_affiliation) %>%
  mutate(political_affiliation = as.factor(political_affiliation)) %>%
  group_by(sub_country_l1_from_source,political_affiliation) %>%
  arrange(desc(totalltems_event_record)) %>%
  summarise(average = mean(totalltems_event_record))





t.test(pol_data$average~pol_data$political_affiliation,var.equal =F)

### Weconclude that there is no relationship b/n the state's ruling party and the amount of garbage present in the beach in that state
```

### Relation b/n county's population and the garbage found in a beach of that county
```{r}
merged_data %>%
  select(sub_country_l2_from_source,totalltems_event_record,population_2019) %>% group_by(sub_country_l2_from_source,population_2019) %>%
  summarise(average = mean(totalltems_event_record) ) %>%
  ggplot(aes(x = population_2019,y = average)) +
  geom_point() +
  geom_text(aes(label = sub_country_l2_from_source),vjust = 1,hjust = 1,check_overlap = T) +
  scale_x_log10(label = comma_format())
  



```


### Preparing the Data for the model

```{r}
merged_data$dow <- as.factor(merged_data$dow)
levels(merged_data$dow)<- list("1" = c("Friday", "Saturday","Sunday"),"0" = c("Monday", "Tuesday","Wednesday","Thursday"))



ch <- times(merged_data$start_time)
merged_data$minute_of_the_day <- 60 * hours(ch) + minutes(ch)


model_data <- merged_data %>%
  select(x,y,total_area_sq_m,totalltems_event_record,political_affiliation,population_2019,dow,minute_of_the_day,month,land_rank) %>%
  mutate(month = as.factor(month),
         political_affiliation = as.factor(political_affiliation),
         land_rank = as.factor(land_rank))


  
```





### Building the model

```{r}
## 75% of the sample size
sample_size <- floor(0.75 * nrow(model_data))

## set the seed to make the partition reproducible
set.seed(321)
train_index <- sample(seq_len(nrow(model_data)), size = sample_size)

train <- model_data[train_index, ]
test <- model_data[-train_index, ]

test <- test %>%
  na.omit()
#model.matrix(~0+., data=model_data) %>% 
  #cor(use="pairwise.complete.obs") %>% 
  #ggcorrplot(show.diag = F, type="lower", lab=TRUE, lab_size=2)
model_1 <- lm(totalltems_event_record ~ .,data = train)
summary(model_1)
```

### Correlation PLot
```{r}
attach(train)
cor(cbind(totalltems_event_record,dow,total_area_sq_m,minute_of_the_day,month,population_2019,political_affiliation))
```
### Residual analysis

```{r}
## Residual Vs Fitted
plot(model_1,which = c(1))

## Residual VS Predictors
plot(head(train$x,1539),rstandard(model_1),main = "Predicted Vs x",xlab = "x") + abline(a = 0,b = 0,col = "red")


plot(head(train$y,1539),rstandard(model_1),main = "Predicted Vs y",xlab = "y") + abline(a = 0,b = 0,col = "red")

## Effect_Plots

effect_plot(model_1,pred = month)
effect_plot(model_1,pred = political_affiliation)
 
## Q-Q plot
model_1_residuals <- rstandard(model_1)
qqnorm(model_1_residuals)
qqline(model_1_residuals)

## Influential points

ols_plot_cooksd_bar(model_1)
shapiro.test(model_1_residuals)

## Predictions
y<- predict(model_1,newdata = test)
y_1 <- test$totalltems_event_record

rmse_1<- sqrt( (y-y_1) %*% (y-y_1) / nrow(test) )
rmse_1

```

### Building another model

```{r}
y <- (1/train$totalltems_event_record)
full_model <- lm(y ~ total_area_sq_m + x+y +population_2019 + political_affiliation + dow,data = train)

base_model <- lm(y ~ total_area_sq_m,data = train)
model_2 <- step(base_model,direction = "forward",scope = formula(full_model))
summary(model_2)
```

### Model 2 Analysis

```{r}
## Residual Vs Fitted
plot(model_2,which = c(1))

## Residual VS Predictors
plot(train$x,rstandard(model_2),main = "Predicted Vs x",xlab = "x") + abline(a = 0,b = 0,col = "red")


plot(train$y,rstandard(model_2),main = "Predicted Vs y",xlab = "y") + abline(a = 0,b = 0,col = "red")

 
## Q-Q plot
model_2_residuals <- rstandard(model_2)
qqnorm(model_2_residuals)
qqline(model_2_residuals)

## Influential points

ols_plot_cooksd_bar(model_2)
shapiro.test(model_2_residuals)

## Predictions
y<- predict(model_2,newdata = test)
y_1 <- test$totalltems_event_record

rmse_2<- sqrt( (y-y_1) %*% (y-y_1) / nrow(test) )
rmse_2
```






